Neural models for computer-assisted musical orchestration: a preliminary study

### 2020-05-08 paper/latex/paper.tex
Add to data gen section that samples are 4 secs

When training a model, the number of samples $n$ to be used in each combination is specified. To create the training data, $n$ instruments are randomly selected from the list of specified instruments, and then a TinySOL sample is randomly selected for each chosen instrument. This gives $n$ different samples, each of a different intrument. These samples are combined to form a soundfile that has each sample playing simultaneously. The melspectrogram is taken of this combination, creating the features to be fed to the model. We created the melspectrogram by \textbf{(Alejandro can you write this of how you generated the melspectrograms using the STFT?)}. We chose to use the melspectrogram because of its ability to represent timbre, a key part of assisted orchestration.

@@

\subsection{Baseline}
In order to have a baseline to compare our results against, we attempted to solve the classification problem using various linear classifiers. The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors. We used the implementations provided in the scikit-learn library for each classifier. For SVM, we used SVC with an RBF kernel. For Random Forest, we set the maximum depth of each tree to be 15. Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem. We found SVM to have the highest accuracy of the three classifiers across all experiments. All of the following baselin experiments used 50,000 generated samples with a train-test split of 60/40. Each sample is a combination of one or more instruments and is four seconds in length. The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.

@@

This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments. However, the linear classifiers were unable to achieve accurate results as the number of instruments increased. Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with linear classifiers.

@@

@@

$$d(x^t, \tilde{x}^t) =\lambda_1 \sum_k \delta_k^t(x_k^t - \tilde{x}_k^t) + \sum_k \delta_k^t|x_k^t - \tilde{x	}_k^t|$$
$$\delta_k^t = 
\begin{cases}
1, \text{if} (x_k - \tilde{x}_k) < 0 \\
0, \text{otherwise}
\end{cases} $$

>>>>

When training a model, the number of samples $n$ to be used in each combination is specified. To create the training data, $n$ instruments are randomly selected from the list of specified instruments, and then a TinySOL sample is randomly selected for each chosen instrument. This gives $n$ different samples, each of a different instrument. These samples are combined to form a soundfile that has each sample playing simultaneously. The audio was trimmed to four seconds in length so our data would all be of the same size. The melspectrogram is taken of this combination, creating the features to be fed to the model. We created the melspectrogram by \textbf{(Alejandro can you write this of how you generated the melspectrograms using the STFT?)}. We chose to use the melspectrogram because of its ability to represent timbre, a key part of assisted orchestration.

@@

\subsection{Baseline}
In order to have a baseline to compare our results against, we attempted to solve the classification problem using various parametric classifiers. The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors. We used the implementations provided in the scikit-learn library for each classifier. For SVM, we used SVC with an RBF kernel. For Random Forest, we set the maximum depth of each tree to be 15. Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem. We found SVM to have the highest accuracy of the three classifiers across all experiments. All of the following baselin experiments used 50,000 generated samples with a train-test split of 60/40. Each sample is a combination of one or more instruments and is four seconds in length. The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.

@@

This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments. However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased. Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with parametric classifiers.

@@

Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length. By changing the mel hop length, which is the distance between the frames of the melspectrogram, an audio signal of any length can be represented as a matrix of the same size as the training data.

\subsection{Evaluation}
-evaluation of orchestrations (not evaluation of model)

@@

$$d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k|$$

$$\delta_{k1} = 
\begin{cases}
1, \text{if   } x_k > \tilde{x}_k \\
0, \text{otherwise}
\end{cases} $$

$$ \delta_{k2} = 1 - \delta_{k1}$$

This metric computes the difference between the FFT of the target and the FFT of the solution.
